Quotients of the magmatic operad:\\ lattice structures and convergent rewrite systems

### 2018-09-11 Mag_quotients.tex
MAJ

We terminate this section with an example illustrating the previous
constructions in the case where the characteristic of the
field $\K$ is different from $2$. For that, we introduce various operads,
which requires the following notations, also used in
Section~\ref{sec:CAs_d}. Let, for any integer $\gamma \geq 1$, the binary
trees $\LComb{\gamma}$ and $\RComb{\gamma}$ be respectively the left and
the right combs of degree $\gamma$. These trees are depicted as
\begin{equation}
    \LComb{\gamma} = \enspace
    \begin{tikzpicture}[xscale=.26,yscale=.3,Centering]
        \node(0)at(0.00,-5.25){};
        \node(2)at(2.00,-5.25){};
        \node(4)at(4.00,-3.50){};
        \node(6)at(6.00,-1.75){};
        \node[NodeST](1)at(1.00,-3.50){\begin{math}\Product\end{math}};
        \node[NodeST](3)at(3.00,-1.75){\begin{math}\Product\end{math}};
        \node[NodeST](5)at(5.00,0.00){\begin{math}\Product\end{math}};
        \draw[Edge](0)--(1);
        \draw[Edge,dotted](1)edge[]node[font=\tiny]{
            \begin{math}\gamma\! -\! 1\end{math}\hspace*{.6cm}}(3);
        \draw[Edge](2)--(1);
        \draw[Edge](3)--(5);
        \draw[Edge](4)--(3);
        \draw[Edge](6)--(5);
        \node(r)at(5.00,1.5){};
        \draw[Edge](r)--(5);
    \end{tikzpicture}
    \qquad \mbox{and} \qquad
    \RComb{\gamma} =
    \begin{tikzpicture}[xscale=.26,yscale=.3,Centering]
        \node(0)at(0.00,-1.75){};
        \node(2)at(2.00,-3.50){};
        \node(4)at(4.00,-5.25){};
        \node(6)at(6.00,-5.25){};
        \node[NodeST](1)at(1.00,0.00){\begin{math}\Product\end{math}};
        \node[NodeST](3)at(3.00,-1.75){\begin{math}\Product\end{math}};
        \node[NodeST](5)at(5.00,-3.50){\begin{math}\Product\end{math}};
        \draw[Edge](0)--(1);
        \draw[Edge](2)--(3);
        \draw[Edge](3)--(1);
        \draw[Edge](4)--(5);
        \draw[Edge,dotted](5)edge[]node[font=\tiny]{
            \hspace*{.6cm}\begin{math}\gamma\! -\! 1\end{math}}(3);
        \draw[Edge](6)--(5);
        \node(r)at(1.00,1.5){};
        \draw[Edge](r)--(1);
    \end{tikzpicture}\,,
\end{equation}
where the values on the dotted edges denote the number of internal
nodes they contain.

We define the \Def{anti-associative operad} by
\begin{equation}
  \label{equ:anti-associative_operad}
  \AAs:=\KMag/_{\IAAs}
\end{equation}
where $\IAAs$ is the ideal spanned by
\begin{math}
  \LComb{3}+\RComb{3}
\end{math}.
\Todo{Illustrer avec la représentation graphique ?}

@@

Finally, for every integer $\gamma\geq 2$, we define the
\Def{$\gamma$-right comb operad} $\RC{\gamma}$ as follows. For every
arity $n$, we let
\begin{equation}
  \label{equ:compositions_of_RC}
  \RC{\gamma}(n):=
    \begin{cases}
        \As(n) &
        \mbox{if } n \leq \gamma, \\
        \RComb{n}
            & \mbox{otherwise},
    \end{cases}
\end{equation}
and the partial composition $\Tfr_1 \circ_i \Tfr_2$ is the partial
composition of $\Tfr_1$ and $\Tfr_2$ in $\Mag$ if the integer
\begin{math}
  n:=|\Tfr_1|+|\Tfr_2| 
\end{math}
is strictly smaller than $\gamma$, and $\RComb{n}$, otherwise.
\medbreak

@@

\medbreak
The lower-bound and the upper-bound of $\As$ and $\AAs$ are given by the
following.
\begin{Theorem}
  \label{thm:example_lattice}
  We have
  \begin{equation}
    \label{equ:example_lower-bound}
    \As\InfQMag\AAs=\NiHil
  \end{equation}
  and
  \begin{equation}
    \label{equ:example_upper-bound}
    \As\SupQMag\AAs=\RC{3}.
    \end{equation}
  \end{Theorem}

\begin{proof}
  First, we show~\eqref{equ:example_lower-bound}. By definition of the
  lower-bound $\InfQMag$, the ideal of relations of $\As\InfQMag\AAs$
  is spanned by the two elements $\LComb{3}-\RComb{3}$ and
  $\LComb{3}+\RComb{3}$. By linear transformations applied to these
  generators, the ideal of relations of $\As\InfQMag\AAs$ is spanned by
  $\LComb{3}$ and $\RComb{3}$, that is it is equal to $\INiHil$, which
  proves~\eqref{equ:example_lower-bound}.

\smallbreak
Now, we show~\eqref{equ:example_upper-bound}. We denote by
\begin{math}
  \pi:\KMag\to\RC{3}
\end{math}
the natural projection. Let $\Tfr$ be a tree of arity $3$. Then,
$\Tfr-\RComb{\gamma}$ belongs to the ideals $\IAs$ and $\IAAs$ of
relations of $\As$ and $\AAs$, respectively. Indeed, every
tree of arity $3$ to the same element $\RComb{\gamma}$. Hence, by
definition of the upper-bound $\SupQMag$, $\pi$ induces a surjective
operad morphism
\begin{math}
  \overline{\pi}:\As\SupQMag\AAs\to\RC{3}
\end{math}.
We conclude using a dimension argument: by definition, the Hilbert series
of $\RC{3}$ is equal to
\begin{equation}
  \label{equ:Hilbert_series_of_RC3}
  t+t^2+2t^3+\sum_{n\geq 4}t^n
\end{equation}
and from the Grassmann formula analog given in
Theorem~\ref{thm:Grassmann_formula_for_Hilbert_series_of_QMag}, we check
that the Hilbert series of $\As\SupQMag\AAs$ is also equal
to~\eqref{equ:Hilbert_series_of_RC3}. Hence, $\overline{\pi}$ is an
isomorphism, which proves~\eqref{equ:example_upper-bound}.
  \end{proof}
\medbreak

>>>>

We terminate this section with an example illustrating the lattice
constructions on $\QMag$ in the case where the
characteristic of the field $\K$ is different from $2$. For that, we
introduce various operads, which requires the following notations, also
used in Section~\ref{sec:CAs_d}. Let, for any integer $\gamma \geq 1$,
the binary trees $\LComb{\gamma}$ and $\RComb{\gamma}$ be respectively
the left and the right combs of degree $\gamma$. These trees are depicted
as
\begin{equation}
    \LComb{\gamma} = \enspace
    \begin{tikzpicture}[xscale=.26,yscale=.3,Centering]
        \node(0)at(0.00,-5.25){};
        \node(2)at(2.00,-5.25){};
        \node(4)at(4.00,-3.50){};
        \node(6)at(6.00,-1.75){};
        \node[NodeST](1)at(1.00,-3.50){\begin{math}\Product\end{math}};
        \node[NodeST](3)at(3.00,-1.75){\begin{math}\Product\end{math}};
        \node[NodeST](5)at(5.00,0.00){\begin{math}\Product\end{math}};
        \draw[Edge](0)--(1);
        \draw[Edge,dotted](1)edge[]node[font=\tiny]{
            \begin{math}\gamma\! -\! 1\end{math}\hspace*{.6cm}}(3);
        \draw[Edge](2)--(1);
        \draw[Edge](3)--(5);
        \draw[Edge](4)--(3);
        \draw[Edge](6)--(5);
        \node(r)at(5.00,1.5){};
        \draw[Edge](r)--(5);
    \end{tikzpicture}
    \qquad \mbox{and} \qquad
    \RComb{\gamma} =
    \begin{tikzpicture}[xscale=.26,yscale=.3,Centering]
        \node(0)at(0.00,-1.75){};
        \node(2)at(2.00,-3.50){};
        \node(4)at(4.00,-5.25){};
        \node(6)at(6.00,-5.25){};
        \node[NodeST](1)at(1.00,0.00){\begin{math}\Product\end{math}};
        \node[NodeST](3)at(3.00,-1.75){\begin{math}\Product\end{math}};
        \node[NodeST](5)at(5.00,-3.50){\begin{math}\Product\end{math}};
        \draw[Edge](0)--(1);
        \draw[Edge](2)--(3);
        \draw[Edge](3)--(1);
        \draw[Edge](4)--(5);
        \draw[Edge,dotted](5)edge[]node[font=\tiny]{
            \hspace*{.6cm}\begin{math}\gamma\! -\! 1\end{math}}(3);
        \draw[Edge](6)--(5);
        \node(r)at(1.00,1.5){};
        \draw[Edge](r)--(1);
    \end{tikzpicture}\,,
\end{equation}
where the values on the dotted edges denote the number of internal
nodes they contain.

We define the \Def{anti-associative operad} by
\begin{equation}
  \label{equ:anti-associative_operad}
  \AAs:=\KMag/_{\IAAs}
\end{equation}
where $\IAAs$ is the ideal spanned by
\begin{math}
  \LComb{3}+\RComb{3}
\end{math}.
\Todo{Illustrer avec la représentation graphique ?}
\medbreak

\begin{Lemma}
\label{lem:properties_of_AAs}
\end{Lemma}

\begin{proof}
\Todo{todo}
\end{proof}
\medbreak

@@

Finally, for every integer $\gamma\geq 2$, we define the
\Def{$\gamma$-right comb operad} $\RC{\gamma}$ as follows. For every
arity $n$, we let
\begin{equation}
  \label{equ:compositions_of_RC}
  \RC{\gamma}(n):=
    \begin{cases}
        \KMag(n) &
        \mbox{if } n \leq \gamma, \\
        \K\Angle{\RComb{n}}
            & \mbox{otherwise},
    \end{cases}
\end{equation}
and the partial composition $\Tfr_1 \circ_i \Tfr_2$ is the partial
composition of $\Tfr_1$ and $\Tfr_2$ in $\Mag$ if the integer
\begin{math}
  n:=|\Tfr_1|+|\Tfr_2| 
\end{math}
is smaller than or equal to $\gamma$, and $\RComb{n}$, otherwise.
\medbreak

@@

\medbreak
The lower-bound and the upper-bound of $\As$ and $\AAs$ in the lattice
$\LatQMag$ are given by the following.
\begin{Theorem}
  \label{thm:example_lattice}
  We have
  \begin{equation}
    \label{equ:example_lower-bound}
    \KAs\InfQMag\AAs=\NiHil
  \end{equation}
  and
  \begin{equation}
    \label{equ:example_upper-bound}
    \KAs\SupQMag\AAs=\RC{3}.
    \end{equation}
  \end{Theorem}

\begin{proof}
  First, we show~\eqref{equ:example_lower-bound}. By definition of the
  lower-bound $\InfQMag$, the ideal of relations of
  $\KAs\InfQMag\AAs$ is spanned by the two elements
  $\LComb{3}-\RComb{3}$ and $\LComb{3}+\RComb{3}$. By linear
  transformations applied to these generators, the ideal of relations of
  $\KAs\InfQMag\AAs$ is spanned by $\LComb{3}$ and $\RComb{3}$,
  that is it is equal to $\INiHil$, which
  proves~\eqref{equ:example_lower-bound}.

\smallbreak
Now, we show~\eqref{equ:example_upper-bound}. We denote by
\begin{math}
  \pi:\KMag\to\RC{3}
\end{math}
the natural projection. Let $\Tfr$ be a tree of arity $3$. From
Lemma~\ref{lem:presentation_of_RC}, the kernel of $\pi$ is spanned by
the elements $\alpha_{\Tfr}:=\Tfr-\RComb{3}$, where $\Tfr$ is a binary
tree of arity $4$. The elements $\alpha_{\Tfr}$ belong to the ideal
$\IAs$ of relations of $\KAs$ and to the ideal $\IAAs$ since both
$[\Tfr]_{\IAAs}$ and $[\RComb{3}]_{\IAAs}$ are equal to $[0]_{\IAAs}$
from Lemma~\ref{lem:properties_of_AAs}. Hence, the kernel of $\pi$ is
included in $\IAs\cap\IAAs$, so that, by definition of the upper-bound
$\SupQMag$, $\pi$ induces a surjective operad morphism
\begin{math}
  \overline{\pi}:\As\SupQMag\AAs\to\RC{3}
\end{math}.
We conclude using a dimension argument: by definition, the Hilbert series
of $\RC{3}$ is equal to
\begin{equation}
  \label{equ:Hilbert_series_of_RC3}
  t+t^2+2t^3+\sum_{n\geq 4}t^n
\end{equation}
and from the Grassmann formula analog given in
Theorem~\ref{thm:Grassmann_formula_for_Hilbert_series_of_QMag}, we check
that the Hilbert series of $\As\SupQMag\AAs$ is also equal
to~\eqref{equ:Hilbert_series_of_RC3}. Hence, $\overline{\pi}$ is an
isomorphism, which proves~\eqref{equ:example_upper-bound}.
  \end{proof}
\medbreak
